Comprehensive Algorithmic Resilience for \\Numeric Applications

### 2013-10-19 paper.tex
Formatting

\section{Introduction}
\label{sec:intro}

@@

\section{Soft Faults}
\label{sec:soft_faults}

@@

\section{Target Applications}
\label{sec:apps}

@@

\subsection{Alternating Direction Method of Multipliers (ADDR)}
\label{sec:apps:lasso}
ADDR is an implementation of an iterative method for solving under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing some function $f(x)$.
It represents the linear solver application domain.
It uses 64-bit precision and spends most of its time in the following linear algebra operations from the GNU Scientific Library~\cite{gsl:2011}: matrix-matrix multiplication (MMM), matrix-vector multiplication (MVM), rank-k update (RK) and Cholesky decomposition (CD).
We use ADDR to solve Lasso problems, where the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$ is minimized.
It is parallelized using MPI and our experiments focus on 4-processors runs with linear systems of size $\{800, 1200, ..., 3600\} \times 500$ as input (denoted $S800$, ..., $S3600$ below).%(800 to 3600 measurements with 500 variables each) and in the experiments below are denoted $S800$, $S1200$, ..., $S3600$.
The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0. \greg{What's the $\sigma$?}
According to our experiments, the running time of this solver is not dependent on the distribution the inputs are sampled from.

\subsection{DRC: Digital Room Correction}
\label{sec:apps:drc}

@@

\subsection{Hattrick}
\label{sec:apps:hattrick}
Hattrick is a sequential application that simulates the motion of bodies under the effects of gravity to help discover extra-solar planets by inferring their existence from Transit Timing Variations.
%, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
It represents the n-body simulation application domain.
Hattrick uses 64-bit precision and spends most of its execution time in the GSL solver for Ordinary Differential Equations to solve the system's equations of motion using the Runge-Kutta method (RK).
A given input is described using three parameters: $P$=number of planets, $T$=amount of time to simulate, and $A$=the accuracy target.
In our experiments we considered the following four inputs: $P2T2090A15$, $P2T3090A15$, $P2T4090A15$ and $P3T2090A11$, where an $A15$ and $A11$ denote accuracy targets of $1e-15$ and $1e-11$, respectively.
\greg{Does it take transit timings as inputs? What are you perturbing?}

@@

\section{Resilience Techniques}
\label{sec:res_tech}

@@

\subsection{Error Detection}
\label{sec:res_tech:err_det}

\vspace{-5pt}
\subsubsection{Algorithmic Detection}
\label{sec:res_tech:err_det:algo}

@@

\subsection{Detection and Tolerance of Pointer Corruptions}
\label{sec:res_tech:pointers}
In 64-bit address spaces corruptions of pointers almost always cause them to refer to un-allocated addresses.
Any accesses via such pointers will be detected by the hardware and result in a Segmentation Fault or Bus Error.
In applications that uses relatively few pointers we detect pointer errors by using these hardware checks and tolerate them via checkpoint-restart.
This is done for ADDR and DRC.
Applications that use pointer-based data structures may experience frequent rollbacks due to pointer errors.
The cost of these rollbacks can be reduced by triplicating the pointers, which is done for Hattrick.
The different pointer replicas can be checked on either some or all accesses.
The former takes less time while the latter has a shorter detection latency.
Every time an error is detected it is immediately corrected, with a rollback required only in the very rare case where multiple replicas are corrupted.

\subsection{Recovery via Checkpoint-Restart and \\Replication}
\label{sec:res_tech:cr}
We tolerate detected but uncorrectable errors via checkpoint-restart.
Each routine's state is recorded at key locations and we roll back to that execution point whenever an error is detected.
Rollbacks are implemented via \texttt{sigsetjmp} and \texttt{siglongjmp}. %\cite{sigsetjmp:1997} routines.
\texttt{sigsetjmp} records the application's execution state (registers and stack pointer) at the time of the checkpoint.
\texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
This light-weight method can roll back to checkpoints recorded in the same function or in a caller function.
When execution returns to the \texttt{sigsetjmp} call the checkpointed portions of the routine's state are recovered and execution repeats.
Pointers and scalars within checkpoints are protected by duplicating them.
Arrays of floating point numbers are protected via a block checksum-based error correction.
Each array is treated as a matrix and divided into blocks of $N \times N$ numbers.
A checksum is computed for each block's rows and columns and used to detect and correct any corrupted entries.
(\greg{Is this a configurable parameter in our experiments?})

@@

\subsection{Evaluation of Subroutine Resilience}
\label{sec:res_tech:eval}

@@

We conducted a comprehensive fault injection campaign where we employed the above resilience techniques with each routine, as applicable.
Fault injection rates between 1e-11 and 1e-8 errors per clock cycle were used\greg{(is this correct)} to probe the effect of one or multiple errors per routine execution.
Our experiments span a wide range of input sizes and a broad range of resilience mechanism configurations, with 100 (\greg{???}) runs for each combination of routine, input, fault rate and resilience configuration.
Table~\ref{tbl:configs} specifies the techniques applied to each routine and its configuration space.
\begin{table}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    \multicolumn{2}{|c|}{Routine}          & Algorithmic Detector             & Checkpointing         & Pointer Replication  \\
    \hline
    \multirow{4}{*}{ADDR}      & MM   & Linear encoding                     & Inputs                & No \\
                               & SYRK & Thresholds:                         & Inputs                & No \\
                               & MVM  & 1e-4 to 1e-7                        & Inputs                & No \\
                               & CD   &                                     & Inputs                & No \\
    \hline
    \multirow{2}{*}{FRC}       & FFT  & Parseval's theorem.                 & Inputs                & No \\
                               & FIR  & Sum conservation.                   & Inputs                & No \\
                               &      & Thresholds: 1e-6 to 3e-8.           &                            \\
    \hline
    \multirow{5}{*}{Hattrick}  & \multirow{5}{*}{RK} & \multirow{5}{*}{Variable step-size} & Periodic Timesteps    & None  \\
                               &      &                                   & Period: 1, 1e3, 1e5   & All pointers, \\
                               &      &                                   &                       & checked at one code location \\
                               &      &                                   &                       & All pointers, \\
                               &      &                                   &                       & checked on each use \\
    \hline
  \end{tabular}
  \caption{The resilience techniques applied to each major routine of each application.}
  \label{tbl:configs}
\end{table}

@@

\subsubsection{Performance Impact}
\label{sec:res_tech:eval:perf}

@@

\subsubsection{Result Accuracy}
\label{sec:res_tech:eval:accuracy}

@@

\section{Evaluation of Resilience Methodology in Applications}
\label{sec:eval}

@@

\subsection{Extrapolation Model}
\label{sec:eval:extrapolation}

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=1.00\columnwidth]{figs/extrapolates.png}
\caption{Extrapolation models}
\label{fig:extrapolations}
\end{figure}

@@

\subsection{Intuitive Accuracy Metric}
\label{sec:eval:errorsensitivity}

@@

% Configurations are described twice.
\subsection{Performance Impact}
\label{sec:eval:perf}
Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time \sui{of parallel (in dashed lines) and sequential (in solid lines) extrapolated instances}, relative to the execution time of the non-fault tolerant version of the extrapolation of the application when no errors are injected.

@@

\subsection{Result Accuracy}
\label{sec:eval:acc}

@@

\subsection{Choice of Optimal Fault Tolerance Techniques}
\label{sec:eval:optchoice}

@@

\section{Summary}
\label{sec:summary}

@@

\bibliographystyle{abbrv}
\bibliography{bibs}

\end{document}

>>>>

\vspace{-10pt}
\section{Introduction}
\vspace{-10pt}
\label{sec:intro}

@@

\vspace{-10pt}
\section{Soft Faults}
\vspace{-10pt}
\label{sec:soft_faults}

@@

\vspace{-10pt}
\section{Target Applications}
\vspace{-10pt}
\label{sec:apps}

@@

\vspace{-10pt}
\subsection{Alternating Direction Method of Multipliers (ADDR)}
\vspace{-10pt}
\label{sec:apps:lasso}
ADDR is an implementation of an iterative method for solving under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing some function $f(x)$.
It represents the linear solver application domain.
It uses 64-bit precision and spends most of its time in the following linear algebra operations from the GNU Scientific Library~\cite{gsl:2011}: matrix-matrix multiplication (MMM), matrix-vector multiplication (MVM), rank-k update (RK) and Cholesky decomposition (CD).
We use ADDR to solve Lasso problems, where the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$ is minimized.
It is parallelized using MPI and our experiments focus on 4-processors runs with linear systems of size $\{800, 1200, ..., 3600\} \times 500$ as input (denoted $S800$, ..., $S3600$ below).%(800 to 3600 measurements with 500 variables each) and in the experiments below are denoted $S800$, $S1200$, ..., $S3600$.
The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0. \greg{What's the $\sigma$?}
According to our experiments, the running time of this solver is not dependent on the distribution the inputs are sampled from.

\vspace{-10pt}
\subsection{DRC: Digital Room Correction}
\vspace{-10pt}
\label{sec:apps:drc}

@@

\vspace{-10pt}
\subsection{Hattrick}
\vspace{-10pt}
\label{sec:apps:hattrick}
Hattrick is a sequential application that simulates the motion of bodies under the effects of gravity to help discover extra-solar planets by inferring their existence from Transit Timing Variations.
%, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
It represents the n-body simulation application domain.
Hattrick uses 64-bit precision and spends most of its execution time in the GSL solver for Ordinary Differential Equations to solve the system's equations of motion using the Runge-Kutta method (RK).
A given input is described using three parameters: $P$=number of planets, $T$=amount of time to simulate, and $A$=the accuracy target.
In our experiments we considered the following four inputs: $P2T2090A15$, $P2T3090A15$, $P2T4090A15$ and $P3T2090A11$, where an $A15$ and $A11$ denote accuracy targets of $1e-15$ and $1e-11$, respectively.
\greg{Does it take transit timings as inputs? What are you perturbing?}

@@

\vspace{-10pt}
\section{Resilience Techniques}
\vspace{-10pt}
\label{sec:res_tech}

@@

\vspace{-10pt}
\subsection{Error Detection}
\vspace{-10pt}
\label{sec:res_tech:err_det}

\subsubsection{Algorithmic Detection}
\vspace{-10pt}
\label{sec:res_tech:err_det:algo}

@@

\vspace{-10pt}
\subsection{Detection and Tolerance of Pointer Corruptions}
\vspace{-10pt}
\label{sec:res_tech:pointers}
In 64-bit address spaces corruptions of pointers almost always cause them to refer to un-allocated addresses.
Any accesses via such pointers will be detected by the hardware and result in a Segmentation Fault or Bus Error.
In applications that uses relatively few pointers we detect pointer errors by using these hardware checks and tolerate them via checkpoint-restart.
This is done for ADDR and DRC.
Applications that use pointer-based data structures may experience frequent rollbacks due to pointer errors.
The cost of these rollbacks can be reduced by triplicating the pointers, which is done for Hattrick.
The different pointer replicas can be checked on either some or all accesses.
The former takes less time while the latter has a shorter detection latency.
Every time an error is detected it is immediately corrected, with a rollback required only in the very rare case where multiple replicas are corrupted.

\vspace{-10pt}
\subsection{Recovery via Checkpoint-Restart and \\Replication}
\vspace{-10pt}
\label{sec:res_tech:cr}
We tolerate detected but uncorrectable errors via checkpoint-restart.
Each routine's state is recorded at key locations and we roll back to that execution point whenever an error is detected.
Rollbacks are implemented via \texttt{sigsetjmp} and \texttt{siglongjmp}. %\cite{sigsetjmp:1997} routines.
\texttt{sigsetjmp} records the application's execution state (registers and stack pointer) at the time of the checkpoint.
\texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
This light-weight method can roll back to checkpoints recorded in the same function or in a caller function.
When execution returns to the \texttt{sigsetjmp} call the checkpointed portions of the routine's state are recovered and execution repeats.
Pointers and scalars within checkpoints are protected by duplicating them.
Arrays of floating point numbers are protected via a block checksum-based error correction.
Each array is treated as a matrix and divided into blocks of $N \times N$ numbers.
A checksum is computed for each block's rows and columns and used to detect and correct any corrupted entries.
(\greg{Is this a configurable parameter in our experiments?})

@@

\vspace{-10pt}
\subsection{Evaluation of Subroutine Resilience}
\vspace{-10pt}
\label{sec:res_tech:eval}

@@

We conducted a comprehensive fault injection campaign where we employed the above resilience techniques with each routine, as applicable.
Fault injection rates between 1e-11 and 1e-8 errors per clock cycle were used\greg{(is this correct)} to probe the effect of one or multiple errors per routine execution.
Our experiments span a wide range of input sizes and a broad range of resilience mechanism configurations, with 100 (\greg{???}) runs for each combination of routine, input, fault rate and resilience configuration.
Table~\ref{tbl:configs} specifies the techniques applied to each routine and its configuration space.
\begin{table}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    \multicolumn{2}{|c|}{Routine}          & Algorithmic                    & Checkpointing           & Pointer \\
    \multicolumn{2}{|c|}{}                 & Detector                       & Checkpointing           & Replication  \\
    \hline
    \multirow{4}{*}{ADDR}      & MM   & Linear encoding                     & \multirow{4}{*}{Inputs} & \multirow{4}{*}{No} \\
                               & SYRK & Thresholds:                         &                         &  \\
                               & MVM  & 1e-4 to 1e-7                        &                         &  \\
                               & CD   &                                     &                         &  \\
    \hline
    \multirow{2}{*}{FRC}       & FFT  & Parseval's theorem.                 & \multirow{4}{*}{Inputs} & \multirow{4}{*}{No} \\
                               & FIR  & Sum conservation.                   &                         &   \\
                               &      & Thresholds:                         &                         &   \\
                               &      & 1e-6 to 3e-8.                       &                         &   \\
    \hline
    \multirow{5}{*}{Hattrick}  & \multirow{5}{*}{RK} & \multirow{5}{*}{Variable step-size} & Periodic & \multicolumn{1}{|l|}{1. None}  \\
                               &      &                                   & Timesteps      & \multicolumn{1}{|l|}{2. All pointers,} \\
                               &      &                                   & Period:        & \multicolumn{1}{|l|}{checked at}  \\
                               &      &                                   & 1, 1e3, 1e5    & \multicolumn{1}{|l|}{one code location}\\
                               &      &                                   &                & \multicolumn{1}{|l|}{3. All pointers,} \\
                               &      &                                   &                & \multicolumn{1}{|l|}{checked on each use} \\
    \hline
  \end{tabular}
  \vspace{-10pt}
  \caption{The resilience techniques applied to each major routine of each application.}
  \vspace{-15pt}
  \label{tbl:configs}
\end{table}

@@

\vspace{-10pt}
\subsubsection{Performance Impact}
\vspace{-10pt}
\label{sec:res_tech:eval:perf}

@@

\vspace{-10pt}
\subsubsection{Result Accuracy}
\vspace{-10pt}
\label{sec:res_tech:eval:accuracy}

@@

\vspace{-10pt}
\section{Evaluation of Resilience Methodology in Applications}
\vspace{-10pt}
\label{sec:eval}

@@

\vspace{-10pt}
\subsection{Extrapolation Model}
\vspace{-10pt}
\label{sec:eval:extrapolation}

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=1.00\columnwidth]{figs/extrapolates.pdf}
\caption{Extrapolation models}
\label{fig:extrapolations}
\end{figure}

@@

\vspace{-10pt}
\subsection{Intuitive Accuracy Metric}
\vspace{-10pt}
\label{sec:eval:errorsensitivity}

@@

% Configurations are described twice.
\vspace{-10pt}
\subsection{Performance Impact}
\vspace{-10pt}
\label{sec:eval:perf}
Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time \sui{of parallel (in dashed lines) and sequential (in solid lines) extrapolated instances}, relative to the execution time of the non-fault tolerant version of the extrapolation of the application when no errors are injected.

@@

\vspace{-10pt}
\subsection{Result Accuracy}
\vspace{-10pt}
\label{sec:eval:acc}

@@

\vspace{-10pt}
\subsection{Choice of Optimal Fault Tolerance Techniques}
\vspace{-10pt}
\label{sec:eval:optchoice}

@@

\vspace{-10pt}
\section{Summary}
\vspace{-10pt}
\label{sec:summary}

@@

\bibliographystyle{abbrv}
\bibliography{bibs}
\end{document}
